---
title: "Peer Graded Assignment 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Applicaions

```{r}
  library(dplyr)
  library(ggplot2)
```

## Loading and Preprocessing the Data

```{r}
dat <- read.csv("./activity.csv", sep = ",", header = TRUE)
dat$date <- as.Date(dat$date)
```

## Mean Total Number of Steps Taken Per Day

I calculated the sum of all the steps taken and then created a histogram showing the total steps taken by day. The average and mediam were also calculated.

```{r}
total_steps <- summarise(group_by(filter(dat,!is.na(steps)), date),
 steps = sum(steps,na.rm = TRUE))

hist(total_steps$steps, breaks = 20, xlab = "Steps", main = "Total Steps/Day")

average_steps <- mean(total_steps$steps, na.rm = TRUE)
median_of_steps <- median(total_steps$steps, na.rm = TRUE)

average_steps
median_of_steps

```

## Average Daily Activity Pattern

Similarly to the previous question, I used the summarise function under the dplyr package. I calculated the averge daily activity pattern and then used ggplot2 to create the time series plot.

```{r}
pattern <- summarise(group_by(filter(dat,!is.na(steps)), interval), 
                     steps = mean(steps, na.rm = TRUE))

ggplot(pattern, aes(x=interval, y=steps), xlab = "Interval", ylab="Steps") +
                    geom_line() + xlab("Interval") + ylab("Steps") +
                    ggtitle("Average Number of Steps")

maximum_steps <- max(pattern$steps)
interval <- pattern[pattern$steps==max(pattern$steps),1]

maximum_steps
interval
```

## Inputting Missing Values

I calculated the amount of missing values. Then I created a separate data set for the missing values. I filled the missing values with the avearage value to that interval and merges the two data sets to calculate the new average and median.

```{r}
new_dat <- dat
sum(is.na(new_dat$steps))
empty_dat  <- is.na(dat$steps)

avg_steps_per_int <- tapply(new_dat$steps, new_dat$interval, mean, na.rm = TRUE)

new_dat$steps[empty_dat] <- avg_steps_per_int[as.character(new_dat$interval[empty_dat])]

total_steps1 <- summarise(group_by(new_dat, date), steps = sum(steps))

hist(total_steps1$steps, breaks = 20, xlab = "Steps", main = "Total Steps/Day")

avg_steps <- mean(total_steps1$steps)
median_steps <- median(total_steps1$steps)

avg_steps
median_steps
```

## Comparing Weekdays to Weekends

I went back to the original data set including the missing values to compare weekedays to weekends. I used mutate and summarise functions to perform the calculations. I then used the ggplot package to visualize the findings.

```{r}
new_dat1 <- mutate(dat, Weektype = ifelse(weekdays(dat$date) == "Saturday"| weekdays(dat$date) == "Sunday", "Weekend", "Weekday"))

pattern1 <- summarise(group_by(filter(new_dat1,!is.na(steps)), interval, Weektype), steps = mean(steps, na.rm = TRUE))

ggplot(pattern1, aes(x = interval , y = steps, color = Weektype)) + 
       geom_line() +
       facet_wrap(~ Weektype, ncol = 1, nrow=2)
```